---
title: "obisindicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{obisindicators}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE, 
  message  = FALSE)
```

```{r setup}
# library(obisindicators)
devtools::document() # update DESCRIPTION, NAMESPACE based on roxygen comments
devtools::load_all() # load the local version of obisindicators
library(dplyr)
```

## Get biological occurrences

Use the 1 million records subsampled from the full OBIS dataset otherwise available at https://obis.org/data/access.

```{r occ}
occ <- obisindicators::occ_1M  # global 1M records subset 
```


## Create an H3 hexagonal grid

```{r}
hex_res <- 1  # hex_res 0 is too big to work, all others work
hex <- obisindicators::make_hex_res(hex_res)
# mapview::mapview(hex)  # show the hex grid with h3 IDs 
# === Then assign cell numbers to the occurrence data:
occ <- occ %>% 
  mutate(
    cell = h3::geo_to_h3(
      data.frame(decimalLatitude, decimalLongitude),
      res = hex_res))
```

## Calculate indicators

The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

Perform the calculation on species level data:

```{r calc_indicators}
idx <- calc_indicators(occ)
```

Add cell geometries to the indicators table (`idx`):

```{r dgcellstogrid}
grid <- hex %>% 
  inner_join(
    idx,
    by = c("hexid" = "cell"))
# you can now visualize with:
# plot(grid["es"])
# mapview::mapview(grid["es"])
```

## Plot maps of indicators

Let's look at the resulting indicators in map form.

```{r map_gcs}
obisindicators::gmap_indicator(
  grid, "es", label = "ES(50)",
  crs="+proj=robin +lon_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

## References

* [crazycapivara/h3-r: R bindings for H3, a hierarchical hexagonal geospatial indexing system](https://github.com/crazycapivara/h3-r)
* [Introduction to H3 for R • H3 for R](https://crazycapivara.github.io/h3-r/articles/h3.html)

# `deckgl`

```{r}
remotes::install_github("crazycapivara/deckgl")
librarian::shelf(
  deckgl, htmlwidgets)

## @knitr h3-cluster-layer
data_url <- paste0(
  "https://raw.githubusercontent.com/uber-common/deck.gl-data/",
  "master/website/sf.h3clusters.json")
# sample_data <- jsonlite::fromJSON(data_url, simplifyDataFrame = FALSE)
sample_data <- data_url

properties <- list(
  stroked = TRUE,
  filled = TRUE,
  extruded = FALSE,
  getHexagons = ~hexIds,
  getFillColor = JS("d => [255, (1 - d.mean / 500) * 255, 0]"),
  getLineColor = c(255, 255, 255),
  lineWidthMinPixels = 2,
  getTooltip = ~mean)

deck <- deckgl(zoom = 10.5, pitch = 20) %>%
  add_h3_cluster_layer(
    data = sample_data, properties = properties) %>%
  add_basemap()

if (interactive())
  deck

```


* [Introduction to r-deckgl • deckgl for R](https://crazycapivara.github.io/deckgl/articles/deckgl.html)
* [Add a h3 cluster layer to the deckgl widget — add_h3_cluster_layer • deckgl for R](https://crazycapivara.github.io/deckgl/reference/add_h3_cluster_layer.html)
* [H3: Area Variation / Nick Rabinowitz / Observable](https://observablehq.com/@nrabinowitz/h3-area-variation)
* [Mapbox Utils / Nick Rabinowitz / Observable](https://observablehq.com/@nrabinowitz/mapbox-utils#fixTransmeridian)
