---
title: "obisindicators"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{obisindicators}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE, 
  message  = FALSE)
```

```{r setup}
library(dplyr)
```

## Get biological occurrences

Use the 1 million records subsampled from the full OBIS dataset otherwise available at https://obis.org/data/access.

```{r occ}
occ <- occ_1M # occ_1M OR occ_SAtlantic
```


## Create an H3 hexagonal grid

```{r}
# remotes::install_github("crazycapivara/h3-r")
# devtools::load_all()
h3_dggs <- obisindicators::make_hex_res(2)
```


Then assign cell numbers to the occurrence data:

```{r dgGEO_to_SEQNUM}
occ$cell <- dgGEO_to_SEQNUM(h3_dggs, occ$decimalLongitude, occ$decimalLatitude)[["seqnum"]]
```

## Calculate indicators

The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

Perform the calculation on species level data:

```{r calc_indicators}
idx <- calc_indicators(occ)
```

Add cell geometries to the indicators table (`idx`):

```{r dgcellstogrid}
library(sf)

grid <- dgcellstogrid(h3_dggs, idx$cell) %>% 
  st_wrap_dateline() %>% 
  rename(cell = seqnum) %>% 
  left_join(
    idx,
    by = "cell")
```

## Plot maps of indicators

Let's look at the resulting indicators in map form.

```{r map_gcs}
# ES(50)
obisindicators::gmap_indicator(grid, "es", label = "ES(50)")
```
