---
title: "Serving OBIS indicators as H3 hexagon tiles (h3t)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Serving OBIS indicators as H3 hexagon tiles (h3t)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE)  # most chunks hit S3 / the deployed h3t service
```

This article shows how to serve [obisindicators](https://github.com/marinebon/obisindicators)
biodiversity metrics --- ES50, Shannon, species richness, number of records ---
as zoomable **H3 hexagon map tiles**, computed on demand from a read-only SQL
query against a cloud-native [DuckDB](https://duckdb.org) store. It realizes the
"biodiversity by hex (OBIS)" pattern of Best et al. (§4.4): *a map tile is just a
viewport-sized projection of a read-only query against an authoritative store.*

The serving half is the [`h3t` tile factory](https://github.com/CalCOFI/api-h3t-py)
(vendored into the [MST server](https://github.com/MarineSensitivity/server) as
`server/h3t`): it base64-decodes a `SELECT` from the tile URL, validates it is
read-only and projects exactly `cell_id, value, n`, substitutes the `{{res}}`
placeholder with the tile's H3 resolution, and returns
[h3j](https://github.com/INSPIDE/h3j-h3t) JSON for MapLibre
(`mapgl::add_h3t_source()`).

```{r setup}
library(obisindicators)
```

## 1. Build the authoritative store

`build_obis_h3_duckdb()` streams [OBIS open-data](https://github.com/iobis/obis-open-data)
geoparquet via DuckDB `httpfs` (byte-range, no full download), bins occurrences
to H3, and writes two layers:

- `idx_h3` --- precomputed **all-taxa** indicators (ES50, Shannon, Simpson,
  richness, n) for resolutions 1--7 (fast default tile layers).
- `occ_h3` --- **species-level** counts at resolution tiers 3/5/7, for
  on-the-fly **taxon/year-filtered** queries.

The indicator math is the SQL translation of [`calc_indicators()`] (`esn = 50`),
pinned to that R reference by the package tests.

```{r build-region}
# a demo region first (cheap) — South Atlantic, matching occ_SAtlantic
build_obis_h3_duckdb(
  src         = "s3://obis-open-data/occurrence/*.parquet",
  path_duckdb = "/share/data/obis/obis_h3.duckdb",
  region_bbox = c(-69.6008, -60, 20.0091, 0.0751))  # lon_min,lat_min,lon_max,lat_max
```

```{r build-global}
# the full global build (expensive S3 scan); drop region_bbox
build_obis_h3_duckdb(
  src         = "s3://obis-open-data/occurrence/*.parquet",
  path_duckdb = "/share/data/obis/obis_h3_global.duckdb")
```

You can also build a small local store from shipped data to develop against:

```{r build-local}
build_obis_h3_duckdb(occ_SAtlantic, "obis_h3.duckdb")
```

## 2. Deploy the tile service

On the MST server the store is registered with the `h3t` service
(`H3T_DBS=obis:/share/data/obis/obis_h3.duckdb`) and fronted by Varnish + Caddy:

```sh
cd /share/github/MarineSensitivity/server
docker compose up -d --build h3t h3tcache   # -> https://h3t.marinesensitivity.org
```

## 3. Compose tile queries

`obis_h3t_sql()` generates the validated `SELECT`. With no filter it reads the
fast precomputed layer; with a `taxon`/`years` filter it computes the indicator
live from the species store.

```{r sql, eval = TRUE}
library(obisindicators)

# default ES50 (all taxa) — precomputed, fast
obis_h3t_sql("es")

# species richness for seabirds since 2000 — computed on the fly
cat(obis_h3t_sql("sp", taxon = list(class = "Aves"), years = c(2000, NA)))
```

`obis_h3t_url()` base64-encodes the SQL into a `?q=` tile URL (and safely
percent-encodes it, so `+`/`/` in the base64 survive query parsing):

```{r url, eval = TRUE}
obis_h3t_url(
  base_url  = "h3tiles://h3tcache.marinesensitivity.org/h3t/{z}/{x}/{y}.h3t",
  indicator = "es",
  release   = "v20260623")
```

Inspect the value distribution (for color-ramp breaks) at the stats endpoint:

```{r stats}
u <- obis_h3t_url(
  base_url  = "https://h3tcache.marinesensitivity.org/h3t/stats",
  indicator = "es")
jsonlite::fromJSON(u)  # {min, max, p02, p98, n}
```

## 4. Map it with MapLibre

```{r map}
library(mapgl)

tiles_es50 <- obis_h3t_url(
  base_url  = "h3tiles://h3tcache.marinesensitivity.org/h3t/{z}/{x}/{y}.h3t",
  indicator = "es",
  release   = "v20260623")

maplibre(center = c(-25, -30), zoom = 3) |>
  add_h3t_source(id = "obis_es50", tiles = tiles_es50) |>
  add_fill_layer(
    id           = "obis_es50",
    source       = "obis_es50",
    source_layer = "obis_es50",
    fill_color   = interpolate(
      column = "value", values = c(1, 50),
      stops  = c("#440154", "#21908C", "#FDE725")),  # viridis
    fill_opacity = 0.8)
```

Swap `indicator = "sp"` for richness, or add a `taxon`/`years` filter to drive a
dropdown --- each choice rebuilds the SQL encoded in the tile URL, and the same
authoritative store answers every query.
