---
title: "Map Regional Diversity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{regional_diversity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE, 
  message  = FALSE)
```

```{r setup}
library(obisindicators)
library(dplyr)
```

## Get biological occurrences for South Atlantic

```{r occ}
occ <- occ_SAtlantic # occ_1M OR occ_SAtlantic
```

## Create a discrete global grid

Create an ISEA discrete global grid of resolution 9 using the [dggridR](https://github.com/r-barnes/dggridR) package:

```{r dgconstruct}
library(dggridR) # remotes::install_github("r-barnes/dggridR")

dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = 9)
```

Here's on overview of all possible resolutions:

```{r dginfo}
inf <- dginfo(dggs)
```

Then assign cell numbers to the occurrence data:

```{r dgGEO_to_SEQNUM}
occ$cell <- dgGEO_to_SEQNUM(dggs, occ$decimalLongitude, occ$decimalLatitude)[["seqnum"]]
```

## Calculate metrics

The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

Perform the calculation on species level data:

```{r calc_es50}
metrics <- calc_es50(occ)
```

Add cell geometries to the metrics table:

```{r dgcellstogrid}
library(sf)

grid <- dgcellstogrid(dggs, metrics$cell) %>% 
  st_wrap_dateline() %>% 
  rename(cell = seqnum) %>% 
  left_join(
    metrics,
    by = "cell")
```

## Plot maps

Let's look at the resulting metrics in map form.

```{r map_gcs}
gmap_metric(grid, "es", label = "ES(50)")
gmap_metric(grid, "shannon", label = "Shannon index")
gmap_metric(grid, "n", label = "# of records", trans = "log10")
```

