---
title: "Regional Diversity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{regional_diversity}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE, 
  message  = FALSE)
```

```{r setup}
library(obisindicators)
library(dggridR)  # remotes::install_github("r-barnes/dggridR")
library(dplyr)
library(sf)
```

## Get regional biological occurrences

Let's use the regional subset for the South Atlantic from the full OBIS dataset otherwise available at https://obis.org/data/access.

```{r occ}
occ <- occ_SAtlantic # occ_1M OR occ_SAtlantic
```

## Create a discrete global grid

Create an ISEA discrete global grid of resolution 9 using the [dggridR](https://github.com/r-barnes/dggridR) package:

```{r dgconstruct}
dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = 9)
```

Then assign cell numbers to the occurrence data:

```{r dgGEO_to_SEQNUM}
occ$cell <- dgGEO_to_SEQNUM(dggs, occ$decimalLongitude, occ$decimalLatitude)[["seqnum"]]
```

## Calculate indicators

The following function calculates the number of records, species richness, Simpson index, Shannon index, Hurlbert index (n = 50), and Hill numbers for each cell.

Perform the calculation on species level data:

```{r calc_indicators}
idx <- calc_indicators(occ)
```

Add cell geometries to the indicators table (`idx`):

```{r dgcellstogrid}
# rda <- here::here("data-raw/region_diversity_stop-dgcellstogrid.RData")
# save.image(rda)
# load(rda)
grid <- dgcellstogrid(dggs, idx$cell) %>% 
  st_wrap_dateline() %>%
  rename(cell = seqnum) %>%
  left_join(
    idx,
    by = "cell")
```

## Plot maps of indicators

Let's look at the resulting indicators in map form.

```{r map_gcs}
# ES(50)
gmap_indicator(grid, "es", label = "ES(50)")

# Shannon index
gmap_indicator(grid, "shannon", label = "Shannon index")

# Number of records, log10 scale, Robinson projection (default)
gmap_indicator(grid, "n", label = "# of records", trans = "log10")

# Number of records, log10 scale, Geographic projection
gmap_indicator(grid, "n", label = "# of records", trans = "log10", crs=4326)
```

