---
title: "resolution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{resolution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(dplyr)
library(sf)
```

## Get biological occurrences

```{r occ}
occ <- occ_SAtlantic # occ_1M OR occ_SAtlantic
```

## Create function to make grid, calculate metrics, and plot maps for different resolution grid sizes

```{r function}
res_changes <- function(resolution = 2){
  hex_res <- 1  # hex_res 0 is too big to work, all others work
  hex <- obisindicators::make_hex_res(resolution)

  # === Then assign cell numbers to the occurrence data:
  occ <- occ %>% 
    mutate(
      cell = h3::geo_to_h3(
        data.frame(decimalLatitude, decimalLongitude),
        res = resolution))
  idx <- calc_indicators(occ)

  grid <- hex %>% 
    inner_join(
      idx,
      by = c("hexid" = "cell"))

  gmap_indicator(grid, "es", label = "ES(50)")
}
```

## Different Resolutions

Details of H3 resolution differences can be found in the [h3geo docs](https://h3geo.org/docs/core-library/restable/).
Resolutions range from 0 (largest) to 15 (smallest).

Generally, resolution 0 is too big to be useful... or even functional, sometimes.
```{r resone}
res_changes(0)
```

```{r restwo}
res_changes(1)
```

At this resolution the S Atlantic is completely covered, meaning that every hex had enough data to compute the ES(50) diversity metric. 
We can see some basic expected patterns such as:
  * higher diversity near to the coast
  * higher diversity near the equator

```{r resthree}
res_changes(2)
```

A this resolution we see gaps throughout the central South Atlantic.
These hexagons did not have enough occurrence records to calculate the diversity metric.

```{r resfour}
res_changes(3)
```
At this higher resolution, gaps dominate the map.
Only places with relatively dense surveying efforts have enough data to calculate the diversity metric.
Note how the relatively data-poor center has a relatively stark boundary spanning from the southern tip of Africa across.
This boundary is visible in the diversity metric plots of lower resolution in the form of a high-low diversity boundary.
The appearance of this abrupt high-low diversity boundary is likely an artifact of how data-poor the central South Atlantic is.
The ES50 diversity metric will bias data-poor to more-diverse when there is extremely low amounts of data.
It should be noted, however, that this bias is *much* less intense than the data-poor to less-diverse in other diversity metrics.
