---
title: "resolution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{resolution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(obisindicators)
library(dplyr)
library(dggridR) # remotes::install_github("r-barnes/dggridR")
library(sf)
```

## Get biological occurrences

```{r occ}
occ <- occ_SAtlantic # occ_1M OR occ_SAtlantic
```

## Create function to make grid, calculate metrics, and plot maps for different resolution grid sizes

```{r function}
res_changes <- function(resolution = 9){
    dggs <- dgconstruct(projection = "ISEA", topology = "HEXAGON", res = resolution)
    occ$cell <- dgGEO_to_SEQNUM(dggs, occ$decimalLongitude, occ$decimalLatitude)[["seqnum"]]
    idx <- calc_indicators(occ)

  grid <- dgcellstogrid(dggs, idx$cell) %>% 
    st_wrap_dateline() %>% 
    rename(cell = seqnum) %>% 
    left_join(
      idx,
      by = "cell")

  gmap_indicator(grid, "es", label = "ES(50)")
}
```

## Different Resolutions

Here we plot res 4, which has hex areas of ~630,000 sq km
This is the highest resolution we can plot without having gaps in the Central Atlantic Region

```{r}
res_changes(4)
```

Here we plot res 6, which has hex areas of ~70,000 sq km.

We still see some gaps and can begin to see the band of high amounts of sampling towards the bottom
This can show us where more sampling is needed

```{r}
res_changes(6)
```

Here we plot res 9, which has hex areas of ~2600 sq km, the default resolution

```{r}
res_changes(9)
```

Here we plot res 11, which has hex areas of ~288 sq km

We see a large amount of gaps throughout the South Atlantic when we get down to this small area

```{r}
res_changes(11)
```
