---
title: "Children taxa & an effort proxy (SPUE) from the OBIS snapshot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Children taxa & an effort proxy (SPUE) from the OBIS snapshot}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  warning  = FALSE,
  message  = FALSE,
  eval     = FALSE)  # DB/map chunks hit the deployed obis_h3 store & tile service;
                     # the pure SQL-builder chunks below opt back in with eval = TRUE
```

This article shows how to filter OBIS by **any WoRMS taxon, at any rank**, using
only the local OBIS snapshot — no calls to the (heavy) OBIS web services — and
how to turn a higher-order taxon into an **observation-effort proxy** for a
presence-only sightings-per-unit-effort (SPUE) indicator.

It builds on [`vignette("h3t")`](h3t.html): the same `occ_h3` species-level
store, plus a baked **`taxon`** table (the WoRMS taxonomy) that lets us resolve
the descendants of an arbitrary AphiaID.

## Why a `taxon` table?

`occ_h3` carries only the six Darwin Core rank columns (`phylum` … `species`)
plus the species-level `aphiaid`. That is enough to filter on, say,
`class = "Aves"`, but **not** on a rank that isn't a column — e.g. Infraorder
*Cetacea*, or an arbitrary clade referenced only by its AphiaID. To support
those we walk the full WoRMS hierarchy:

- `build_taxon_parquet.R` extracts `taxonID` (= AphiaID), `parentNameUsageID`,
  `acceptedNameUsageID`, `scientificName`, `taxonRank` from a WoRMS DarwinCore
  `taxon.txt` into a compact `taxon.parquet`.
- `migrate_add_taxon.R` bakes that into `obis_h3.duckdb` (indexed on
  `parentNameUsageID`), alongside `occ_h3` / `idx_h3`.

```{r setup, eval = TRUE}
library(obisindicators)
```

```{r connect}
library(DBI)
library(duckdb)

con <- dbConnect(duckdb(), "/share/data/obis/obis_h3.duckdb", read_only = TRUE)
dbExecute(con, "LOAD h3;")
```

## Resolve the children of a taxon

`obis_taxon_children()` returns the seed taxon plus every descendant, at any
rank, via a recursive walk of `parentNameUsageID`. On the full WoRMS table
(~1.5 M rows) this resolves in a fraction of a second.

```{r children}
# Infraorder Cetacea = AphiaID 2688 (a rank that is NOT an occ_h3 column)
cet <- obis_taxon_children(2688, con)
nrow(cet)                              # ~1,665 descendant taxa
table(cet$taxonRank)                   # Species, Genus, Family, ...

# Matt Biddle's phytoplankton example: class Bacillariophyceae (diatoms) = 148899
dia <- obis_taxon_children(148899, con)
nrow(dia)                              # ~71,500 descendant taxa
```

The descendant AphiaIDs (`cet$taxonID`) are exactly the set used to filter
`occ_h3.aphiaid`. `obis_taxon_subtree_sql()` returns that resolution as a
standalone read-only `SELECT`, handy for the API.

## Map a taxon's records (any rank)

`obis_h3t_sql(aphiaid = ...)` builds a served tile query that filters `occ_h3`
to the subtree via a recursive CTE (the `h3t` validator permits
`WITH RECURSIVE`). `obis_h3t_url()` base64-encodes it into a tile URL.

```{r map-records-sql, eval = TRUE}
sql <- obis_h3t_sql(indicator = "n", aphiaid = 2688)   # # records, all Cetacea
cat(sql)

url <- obis_h3t_url(
  base_url  = "h3tiles://h3t.marinesensitivity.org/h3t/{z}/{x}/{y}.h3t",
  sql       = sql)
```

```{r map-records, eval = TRUE}
library(mapgl)
maplibre(center = c(-40, 20), zoom = 1.5) |>
  add_h3t_source(id = "cet", tiles = url) |>
  add_fill_layer(
    id = "cet_fill", source = "cet", source_layer = "cet",
    fill_color = interpolate(
      column = "value", values = c(1, 1000),
      stops  = c("#3b528b", "#5ec962")))
```

## An effort proxy: sightings per unit effort (SPUE)

A raw record map conflates *abundance* with *observation effort* — a hexagon
lights up both where a species is common and where people looked hard. When
records come from multi-species surveys, a **higher-order taxon** gives a
presence-only proxy for that effort: within the footprint of, say, all Cetacea
(the surveys that could have recorded any whale/dolphin), what fraction of
records are the target species?

$$\text{SPUE}_{cell} = \frac{\text{records of the target taxon}}{\text{records of the effort taxon}}$$

`obis_spue_sql()` builds this as a served tile query (two recursive subtrees:
numerator = target, denominator = effort). `calc_spue()` is the pinned R
reference (see `test-spue-parity.R`).

```{r spue-sql, eval = TRUE}
# target: Tursiops truncatus (common bottlenose dolphin) = 137111
# effort: Infraorder Cetacea = 2688 (its multi-species-survey footprint)
spue_sql <- obis_spue_sql(num_aphiaid = 137111, den_aphiaid = 2688)
cat(spue_sql)

spue_url <- obis_h3t_url(
  base_url = "h3tiles://h3t.marinesensitivity.org/h3t/{z}/{x}/{y}.h3t",
  sql      = spue_sql)
```

```{r spue-map, eval = TRUE}
maplibre(center = c(-40, 20), zoom = 1.5) |>
  add_h3t_source(id = "spue", tiles = spue_url) |>
  add_fill_layer(
    id = "spue_fill", source = "spue", source_layer = "spue",
    fill_color = interpolate(
      column = "value", values = c(0, 1),
      stops  = c("#440154", "#fde725")))
```

The tooltip `n` carries the **effort** count (denominator): a bright cell backed
by a large `n` is a robust SPUE; a bright cell over `n = 3` is noise. Choosing
the effort taxon is the analyst's call — the parent class or order is a sensible
default, but a survey-defined group (e.g. Cetacea, seabirds) is often better.
See [`vignette("scaling")`](scaling.html) for how the denominator's sparsity
behaves across H3 resolutions.

```{r disconnect}
dbDisconnect(con, shutdown = TRUE)
```
